Modelling Structured Societies: a Multi-relational Approach to Context Permeability

### 2014-08-18 doc/main.tex
pre-proof-read version of the paper added; some correlation analysis added; other old analysis corrected

\documentclass[preprint,number]{elsarticle}
\usepackage[utf8]{inputenc}
\usepackage{multirow}

@@

\begin{abstract}
The structure of social relations is fundamental for the construction of plausible simulation scenarios. It shapes the way actors interact and create their identity within overlapping social contexts. Each actor interacts in multiple contexts located within relations that constitute their social space. We present a modelling approach to construct structured agent societies with multiple concomitant social networks. We study the notion of context permeability using a consensus game in which agents try to achieve global consensus. We design and analyse models of permeability with three model different configurations. In the first model agents interact concurrently in multiple social networks. The second model uses a context switching mechanism that adds temporal dynamics to agent interaction. They switch between the different networks spending more or less time in each network. Finally we use a strategic neighbourhood selection by segregation that acts as a regulatory mechanism in the societal self-organisation process to achieve consensus. We compare each model and analyse the influence of different social topologies in the speed of convergence to global consensus.
%TODO abstract: brief strong conclusion
\end{abstract}

@@

\end{frontmatter}
\newpage
%********************************************
%				INTRODUCTION
%********************************************
\section{Introduction and Related Work}
\label{sec:introduction}
%Social simulation view oever consensus
\noindent In social simulation, studying models of consensus formation helps us to investigate real-world phenomena. This can be done not only by constructing data-driven simulation but also with the help of abstract models that permit the assembly of \textit{what-if-scenarios}. Examples of real-world target phenomena in this domain include: the joint assessments of social policies or, in the context of economics and politics, \textit{the voting problem}. Herbert Simon investigated this last problem in an early model described in~\cite{Simon1954}.

Formal opinion dynamics models provide an understanding if not an analysis of opinion formation processes.  An early formulation of these models was designed to comprehend complex phenomena found empirically in groups \cite{French1956}. In particular, the work on consensus building in the context of decision-making was first studied by DeGroot \cite{Degroot74} and Lehrer \cite{Lehrer1975}. Empirical studies of opinion formation in large populations have methodological limitations, as such, we use simulation, in particular Multi-Agent Simulation (MAS), as a methodological framework to study such phenomena in a larger scale. Most opinion dynamics simulation models are based either on binary opinions \cite{Galam1997,Antunes2009} or on opinions with continuous values \cite{Deffuant2000,Deffuant2002}. In these models, agents update their opinions either under social influence or according to their own experience. For a detailed analysis over some opinion dynamics model analytical and simulation results, refer to \cite{Hegselmann2002}.

In agent-based opinion dynamics models, agent interactions are guided by social space abstractions. In some models, dimensionality is irrelevant. Typically, all agents can participate in interactions with all other agents. Axelrod, takes a different approach in his model of dissemination of culture \cite{Axelrod1997} and represents agents in a bi-dimensional grid which provides the structure that shapes how they interact. In Weisbuch's bounded confidence model with social networks \cite{Weisbuch2004}, the agents are bound by network structures. A first definition for these types of bounded confidence models was given in \cite{Krause1997}.

%position, most agent-based models don't take into account  social relations and this is important
In real-world scenarios, actors engage in a multitude of social relations different in kind and quality. Most simulation models don't explore social space designs that take into account the differentiation between coexisting social networks. Modelling multiple concomitant relations was an idea pioneered by Peter Albin in \cite{Albin1975} but without significant further development. The complex social structures that result from interacting in these different dimensions form the basis for the formation of social identity \cite{Roccas2002,Ellemers2002}.

%A broader view over consensus formation
\subsection{On Consensus}
Opinion dynamics models try to study under which circumstances convergence to consensus can be achieved. These agent-based models can have broader applications outside social simulation. In distributed systems in general, consensus is a means by which processes agree on some data value that is needed during computation, typically, this agreement is the result of a negotiation process often with the aid of a mediator. In human societies, social conventions emerge to deal with coordination and subsequently with cooperation problems \cite{Lewis1969}, these conventions are regularities of behaviour that can turn normative with time due to their being persistent solutions to such problems. Moreover, conventions typically originate from the necessity to resolve conflict. Mult-Agent Systems (MAS) can also benefit from being capable of produce emergent conventions as they might need to coordinate or cooperate in order to solve different problems \cite{Delgado2002}.

The decentralised nature of these models of consensus is highly desirable for problems where dynamic control is needed. In these scenarios, creating conventions before hand (off-line) or developing a central control mechanism for generating them can be a difficult and intractable task. MAS can be designed to update their values and strategies based on this local information. Consequently, the guarantee of achieving a global consensus and the efficiency of convergence towards it has been a natural concerns in the design of local interaction behaviours. Shoham and Tennenholtz compared several strategy update rules in %paper urbano

%TODO this was the last thing I wrote

. This explains the interest on the design of
emergent conventions through a co-learning process [13]: individual agents occasionally meet and observe each
other, and they may .

%TODO present the literature related to norms and consensus and motivate the problem

%TODO distributed systems need consensus to perform reliable sets of actions (tolerance to faults etc) such as transations
%Agents that co-exist need to coordinate to perform certain tasks or choose appropriate strategies, science a particular behaviour can 
%have an impact on the each agent not only on the environment, and even when agents have an impact on the environment, this is reflected
%in the performance of other agents that co-habit the same environment.

%TODO as in human societies, consensus is not an intentional end result but rather an emergent process which arises 
%without a central authority

%TODO see Emergent Conventions and the Structure of Multi{Agent Systems
%TODO see On the synhesis of useful social laws for artificial agent societies

In the field of Distributed Artificial Intelligence (DAI) in general and in Multi-Agent Systems (MAS) in particular, achieving a global agreement on some strategy, convention or norm is fundamental 
\cite{Delgado2002}

%Objective of this paper
This paper is aimed at extending the line of research regarding the representation of social spaces by modelling these with explicit multiple networks.

%****************************************
%			PAPER STRUCTURE
%****************************************
\subsection{Paper Structure}
This paper is organised as follows. In the next section we define our modelling approach that uses multiple social networks. We describe our game of consensus and introduce multiple model variations designed to study the notion of context permeability and the process of consensus formation. Section \ref{sec:experimental-setup} describes the experimental set-up for our series of experiments. Finally, in section \ref{sec:results-discussion}, we discuss our results and compare the different simulation models.

\section{Experimental Setup}
\label{sec:experimental-setup}

@@

\section*{References}
\bibliographystyle{elsarticle-num}
\bibliography{main}

>>>>

\documentclass[preprint,number]{elsarticle}
\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage{amsthm}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{appendix}
\usepackage{subcaption}
\usepackage{siunitx}

\usepackage{pdflscape}

\usepackage[draft]{todonotes}
\newcommand{\comment}[1]
{\par {\bfseries \color{blue} #1 \par}} %comment showed

%scientific notation setup
\sisetup{
	round-mode = places,
	round-precision = 2,
	scientific-notation = false}

\usepackage{booktabs}

\usepackage{float}
\floatname{algorithm}{Model}
\newcommand{\algorithmname}{Model}

\newtheorem{mydef}{Definition}

@@

		\begin{abstract}
			The structure of social relations is fundamental for the construction of plausible simulation scenarios. It shapes the way actors interact and create their identity within overlapping social contexts. Each actor interacts in multiple contexts located within relations that constitute their social space. We present a modelling approach to construct structured agent societies with multiple concomitant social networks. We study the notion of context permeability using a consensus game in which agents try to achieve global consensus. We design and analyse two different models of permeability. In the first model agents interact concurrently in multiple social networks. The second model uses a context switching mechanism that adds temporal dynamics to agent interaction. They switch between the different networks spending more or less time in each network. We compare each model and analyse the influence of different social networks in the speed of convergence to consensus. We conduct a series of experiments that show the impact of coexisting social networks when designing simulation models. This approach unveils both the limitations of the current modelling approaches and possible research directions for complex social space simulations.
		\end{abstract}

@@

	\end{frontmatter}
	\newpage
	%******************************************
	%		INTRODUCTION
	%******************************************
	\section{Introduction}
	\label{sec:introduction}
	%******************************************
	%		Research context
	%******************************************
	Studying simulation models of consensus formation can advance our understanding of real-world emergent social phenomena. Typically, we can construct models with different levels of abstraction. Simulation methods can range from data-driven paradigms to more abstract descriptions that allow us to create \textit{what-if-scenarios}. While most models are simplified descriptions of the real-world, many constitute complex systems themselves and thus the necessity of tools like simulation to explore their properties. In \cite{Emmeche1997}, Emmeche makes a distinction between two types of complexity. He characterises the complexity of real systems as \textit{``ontological complexity"} and the one of models as \textit{``descriptive complexity"}. Examples of real-world target phenomena for social simulation models include for instance the joint assessments of policies or, in the context of economics and politics, \textit{the voting problem}. Herbert Simon investigations on this problem were also one of the first stepping stones to the field of social simulation \cite{Simon1954}.

	In agent-based \textit{opinion dynamics}, agent interactions are guided by \textit{social space} abstractions. In some models, dimensionality is irrelevant and any agent can interact with any other agents.
	Other approaches use an underlying structure to create agent neighbourhoods. Axelrod for instance, in its model of \textit{dissemination of culture} \cite{Axelrod1997}, represents agents in a bi-dimensional grid. Finally, in an attempt to mimic real social systems, some simulation models make use of \textit{complex network models} (see \cite{Weisbuch2004}) to create the infrastructure that guides agent interaction.

	%******************************************
	%		Motivation 
	%******************************************
	%position, most agent-based models don't take into account  social relations and this is important
	In real-world scenarios, actors engage in a multitude of social relations different in kind and quality. Most simulation models don't explore \textit{social space} designs that take into account the differentiation between coexisting social networks. Modelling multiple concomitant relations was an idea pioneered by Peter Albin in \cite{Albin1975} but without any further development. The complex social structures that results from interacting in these different dimensions forms the basis for the formation of social identity \cite{Roccas2002,Ellemers2002}.

	%******************************************
	%		Objectives
	%******************************************
	In this paper, we explore the possibility of modelling opinion dynamics with multiple social networks. We look at how the properties influence the convergence to opinion consensus. Starting with this premise of 
	using multiple social networks, we present a series of models that use these networks in different ways which creates distinct emergent dynamics. We want to observe the consequences of using multiple social networks at the same time ---while maintaining the model as simple and abstract as possible much like what happens with the current contributions on opinion dynamics present in the literature.

	In our models agents can: interact at the same time in the multiple networks (choosing partners from any network) or switch between networks (choosing only partners from their current network). While our models seem quite simple, and are quite abstract, they generate quite complex outcomes and thus the need of simulation to understand their dynamics.

	%******************************************
	%		Paper Structure
	%******************************************
	\subsection{Paper Structure}
	This paper is organised as follows. In the next section we present work related to, \textit{opinion dynamics}, \textit{social space} modelling and \textit{complex network models}. In the following section, we describe our game of consensus and introduce multiple model variations designed to study the notion of context permeability and the process of consensus formation in multiple social networks. Section \ref{sec:experimental-setup} describes the experimental set-up; the set of tool and methodologies followed to conduct our investigations. In section \ref{sec:results-discussion}, we present and discuss our results and compares the different simulation models. Finally we conclude by making a summary of what we learned from the simulation models and point out possible future research directions.

	%******************************************
	%		Related Work
	%******************************************
	\section{Related Work}
	\label{sec:relatedwork}
	%******************************************
	%	Opinion Dynamics
	%******************************************
	\subsection{Opinion Dynamics and Consensus Formation}
	Formal opinion dynamics models provide an understanding if not an analysis of opinion formation processes.  An early formulation of these models was designed to comprehend complex phenomena found empirically in groups \cite{French1956}. In particular, the work on consensus building in the context of decision-making was first studied by DeGroot \cite{Degroot74} and Lehrer \cite{Lehrer1975}. Empirical studies of opinion formation in large populations have methodological limitations, as such, computational sociology arises with a set of tools to cope with those limitations. We use simulation, in particular \textit{multi-agent simulation (MAS)}, as a methodological framework to study such phenomena in a larger scale. Most opinion dynamics simulation models are based either on binary opinions \cite{Galam1997,Antunes2009} or continuous values \cite{Deffuant2000,Deffuant2002}. In these models, agents update their opinions either under social influence or according to their own experience. For a detailed analysis over some opinion dynamics model analytical and simulation results, refer to \cite{Hegselmann2002}.

	Opinion dynamics models try to study under which circumstances consensus or polarisation is reached. Agent-based models can have broader applications outside social simulation. In distributed systems in general, consensus is a means by which processes agree on some data value that is needed during computation, typically, this agreement is the result of a negotiation process often with the aid of a mediator. In human societies, social conventions emerge to deal with coordination and subsequently with cooperation problems \cite{Lewis1969}, these conventions are regularities of behaviour that can turn normative with time due to their being persistent solutions to such problems. Moreover, conventions typically originate from the necessity to resolve conflict. MAS can also benefit from being capable of produce emergent conventions in coordination or cooperation problems \cite{Delgado2002}. The decentralised nature of these models of consensus is highly desirable for problems where dynamic control is needed. In these scenarios, creating conventions before hand (off-line) or developing a central control mechanism for generating them can be a difficult and intractable task. MAS can be designed to update their values and strategies based on this local information. Consequently, the guarantee of achieving a global consensus and the efficiency of convergence towards it has been a natural concerns in the design of local interaction behaviours.

	Consensus models are also investigated as means to create conventions in MAS. Shoham and Tennenholtz \cite{Shoham1994} create a bridge between economic literature and machine learning by studying a series of models in which multiple agents are engaged in learning a particular convention. They call this process \textit{co-learning}. The complexity of these systems derives from the its concurrent nature: one agents adapts to the behaviour of another agent it has encountered, as these agents update their behaviour in a similar fashion, this results in a highly non-linear system dynamics. In their work, Shoham and Tennenholtz define the notion of \textit{stochastic social games} and compare different rules with the objective of establishing conventions in a decentralised fashion. In their stochastic games, they present one particular rule which we use in our models:

	\begin{mydef}
		\label{def:external-majority}
		\textit{External Majority (EM)} update rule: adopt action $i$ if so far it was observed in other agents more often than other action and remain with your current action in the case of equality. 
	\end{mydef}

	This EM was shown to coincide with \textit{Highest Cumulative Reward (HCR)} which is a simple rule that states that ``an agent should adopt an action that has yielded the highest cumulative reward to date". We chose to use EM in our simulation models for its simplicity and success in the evolution of conventions. We can see that there is an overlap between opinion dynamics from social simulation and agent based learning---or co-learning in this case.

	%******************************************
	%	Social Structure in Simulation Models
	%******************************************
	\subsection{Social Structure in Simulation Models}
	One of the most widely used approaches to represent social spaces in abstract simulation models is using a  bi-dimensional gridd. This approach has its origins in the first \textit{“checkerboard”} model introduced in computational social sciences by Thomas Schelling \cite{Schelling1971} and Sakoda \cite{Sakoda1971}. Another famous example that uses bi-dimensional grids as social structure is the social simulation model of dissemination of culture from Axelrod \cite{Axelrod1997}.

	Cellular automata (CA) is an example from the are of artificial life that also makes use of bi-dimensional structures to model neighbourhood interactions. While these models are idealised frameworks, the exploration of simple computational models can bring us deeper insights than models with high levels descriptive complexity which would render their analysis, very difficult. The work of Flache and Hegselmann \cite{FlacheHegselmann2001} relaxes the standard CA assumptions about the regular bi-dimensional grids by considering irregular grid structures (Voronoi diagrams). This work presents results on the robustness of some important general properties in relation to variations in the typical grid structure. We can also find contributions that incorporate both the complexity of network models and the behaviour of dynamic processes. One example of this combination is the work in \cite{Sullivan2000} which explores a graph-based cellular automaton to study the relationship between spatial form of urban systems and the robustness of different process dynamics under spatial change. Moreover this allows for the integration of real geographical information to construct the graph-cellular automata. This fact makes so that while the models of dynamic processes are relatively abstract, they are tested in spatial structures based on data from real complex structures and thus making a connection between purely abstract models and data-driven models.

	%******************************************
	%	Complex Network Models
	%******************************************

	Finally, another type social space models are the \textit{random graph / network models}. Networks with different properties can be built using network models. Each complex network or class of complex networks, presents specific topological features. These properties are typically found in---and the models inspired by---real-world network structures. One of the first complex network models is the random graph from Erdős and Renyi \cite{Erdos1959}. Other examples include the model of preferential attachment \cite{Barabasi1999} and the small-world network models \cite{Watts1998}. One common problem of these models is the fact that once generated, the network structure is static so these models are not really suitable for the description of highly dynamic groups or communities. An example of a model that takes these dynamics into account is the model of team formation~\cite{Gaston2005}. A more in-depth review of network models and their properties is given in~\cite{Nunes2012}.

	\section{Multi-context Models}
	\label{sec:multi-context-models}
	In this section, we present our modelling approach to explore the concept of permeability between contexts. In our approach, we considers multiple social networks as a way to represent the complex social space of an agent. This setting can be seen in a simulation as a n-dimensional scenario where each dimension surface represents a different social relation (see figure \ref{img:multiple-relations}). Each relation is instantiated using a social network model. The word ``context" is used in many different senses and it is subjected to many analysis \cite{Hayes1997}. Here, we use the term \textit{social context} in a simpler and more strict way. Agents belong to distinct social contexts (neighbourhoods) in these multiple networks. The \textit{social context} of an agent is thus a set of neighbours currently available for interaction. \textit{Context Permeability} is the ability of a particular norm, strategy, opinion or trend to permeate from one social relation to another. Agents belonging to disjoint neighbourhoods in different networks foster the creation of bridges between subsets of the population that wouldn't otherwise be influenced by a consensus formed in a distant coalition or in a distinct context.

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.7\linewidth]{./images/multi-context_model.pdf}
		\begin{minipage}{0.9\linewidth}
			\caption{Multiple social network structure that shapes the social space in our simulation models.}
			\label{img:multiple-relations}
		\end{minipage}
	\end{figure}

	%Consensus
	We study the notion of context permeability in different simulation models in which agents interact using a simple consensus game. The society of agents has to adopt a binary \textquotedblleft option \textquotedblright according to a majority rule. We use the speed of consensus as a measure for self-organisation, exploring the relationship between different network topologies and this measure.

	%Presenting 3 models
	We partitioned the design space of our experiments using different simulation models. Each model is designed to analyse different aspects of social context permeability. In a first model \cite{Antunes2007, Antunes2010}, we study the notion of permeability by overlapping social networks. In a second model \cite{Antunes2009}, we analyse the dynamics introduced by switching between \textit{social contexts}(neighbourhoods in different networks), adding a temporal component to neighbourhoods within different social relations.

	\subsection{Simulation and Consensus Game}
	\label{sec:sim_consensus}
	Our agent-based models are designed as \textit{discrete-event} models. On each simulation cycle, every agent executes a simulation step. The agents are selected in a random uniform fashion for execution. This is common practice in agent-based simulation and guarantees that there is no bias caused by the order in which the agents are executed. For each simulation model, we present a description of the individual agent behaviour. Every model we present can be seen as a binary opinion dynamics model or a consensus game. In this game, the society of agents tries to reach an arbitrary global consensus about two possible choices. Each agent keeps track of the number of choices observed from interaction partners throughout a simulation run. In each iteration, each agent selects an available neighbour to interact with and observes its current choice. The agent decides to switch its current choice if the observed option becomes the majority of the two possible choices. The objective of the game is to reach a global consensus, but the particular choice that gets collectively selected is irrelevant. What is important is that overall agreement is achieved. In the consensus game we consider, each agent uses the previously presented external majority (EM) rule to update its opinion value (see definition \ref{def:external-majority} in section \ref{sec:relatedwork}).

	\newpage
	\subsection{Context Permeability}
	This first model of context permeability \cite{Antunes2007,Antunes2010} is designed to study our first setting for permeability. In most interesting applications, an agent is immersed in a complex social world where it engages in a multitude of relations with other agents (and even aggregate agents such as institutions or informal groups). In this model we consider that the \textit{context permeability} is created by agents that can belong simultaneously to several relations. As an example, two agents can be simultaneously family and co-workers, and while some links might not connect two agents directly, others can contain neighbourhoods in which they are related~(either directly or through a common neighbour). Model \ref{model:context_permeability} describes the behaviour of each agent for each simulated step.
	%***************************************************
	%	Context Permeability Model Description
	%***************************************************
	\begin{algorithm}[H]
		\caption{Context Permeability}
		\label{model:context_permeability}
		\algsetup{indent=2em}
		\begin{algorithmic}
			\vspace{0.5em}
			\STATE $M$ \COMMENT{Simulation Model / Environmnet}
			\STATE $A$ \COMMENT{Current Agent in Execution}
			\\ \hrulefill

			\STATE\COMMENT{\textbf{Randomly select a network}}
			\STATE $r$ $\leftarrow$ randomUniform(0,$ M.networks.length() $)
			\STATE network $\leftarrow$ $M.networks[r]$

			\STATE
			\STATE \COMMENT{\textbf{Randomly select a neighbour}}
			\STATE neighbours $\leftarrow$ network.neighboursOf($A$) 
			\STATE $r$ $\leftarrow$ randomUniform(0,neighbours.length())
			\STATE partner $\leftarrow$ neighbours[r]
			\STATE
			\STATE \COMMENT{\textbf{Update opinion and memory}}
			\STATE \COMMENT{opinions take the values \{0,1\}--so they can be used as an index}
			\STATE A.memory[partner.opinion]++
			\IF {(A.memory[partner.opinion] $>$ A.memory[A.opinion])}
			\STATE A.opinion $\leftarrow$ partner.opinion
			\ENDIF
		\end{algorithmic}
	\end{algorithm}
	In this model, on each time step, each agent selects a random neighbour from any 
	social network available and updates its opinion based on the opinion of the selected partner using the \textit{external majority} rule.

	\subsection{Context Switching}
	\label{sec:models_cs}
	In the context switching model \cite{Antunes2009}, agents are embedded in multiple relations represented as static social networks and they switch \textit{contexts} with some probability $\zeta_{C_i}$ associated with each context $C_i$. The switching probability value is the same in the same network. Each agent switches from its current neighbourhood to a its neighbourhood in another network (in which this agent was inactive). The agents are only active in one context at a time and can only perform encounters with neighbours in the same context. We can think of context switching as a temporary deployment in another place, such as what happens with temporary immigration. The network topology is static; when an agent switches from one network to another, they become inactive in one network and active in their destination. 
	%***************************************************
	%	Context Switching Model Description
	%***************************************************
	\begin{algorithm}[H]
		\caption{Context Switching}
		\label{model:context_switching}
		\algsetup{indent=2em}
		\begin{algorithmic}
			\vspace{0.5em}
			\STATE $M$ \COMMENT{Simulation Model / Environmnet}
			\STATE $A$ \COMMENT{Current Agent in Execution}
			\\ \hrulefill

			\STATE\COMMENT{\textbf{Get the current network (\textit{context} is a network index)}}
			\STATE cNetwork $\leftarrow$ M.networks[A.context]
			\STATE 
			\STATE \COMMENT{\textbf{Get the neighbours for the current network}}
			\STATE neighbours $\leftarrow$ cNetwork.neighboursOf($A$) 
			\STATE \COMMENT {filter by agents active in the same network}
			\STATE neighbours $\leftarrow$ \{n $|$ n $\in$ neighbours $\wedge$ (n.context = A.context)\}
			\STATE
			\STATE\COMMENT{\textbf{Randomly select a neighbour}} 
			\STATE $r$ $\leftarrow$ randomUniform(0,neighbours.length())
			\STATE partner $\leftarrow$ neighbours[r]
			\STATE
			\STATE \COMMENT{\textbf{Update opinion and memory}}
			\STATE \COMMENT{opinions take the values \{0,1\}--so they can be used as an index}
			\STATE A.memory[partner.opinion]++
			\IF {(A.memory[partner.opinion] $>$ A.memory[A.opinion])}
			\STATE A.opinion $\leftarrow$ partner.opinion
			\ENDIF
			\STATE
			\STATE \COMMENT{\textbf{Switch to another network}}
			\STATE switchingProb $\leftarrow$ M.params.switchingProb(A.context)
			\STATE $r \leftarrow $ randomUniform(0,1)
			\IF{($r < $ switchingProb)}
			\STATE numNets $\leftarrow$ M.networks.length()
			\STATE nextNetworks $\leftarrow$ \{i $|$ (i $\in$ [0,numNets[) $\wedge$ (M.networks[i] != A.context)\}
			\STATE $r$ $\leftarrow$ randomUniform(0,nextNetworks.length())
			\STATE A.context $\leftarrow$ nextNetworks[r]
			\ENDIF
		\end{algorithmic}
	\end{algorithm}
	\noindent In this model agents select an (active) partner from their current neighbourhood. If a partner is available the agents update their opinion value based on EM. At the end of the interaction, the agent switches \textit{from} the current context to a different one with a probability associated with its current network.

	This model describes an abstract way to represent the time spent on each network using the switching probability $\zeta_{C_i}$. Here, the permeability between contexts is achieved using this temporal component. Context switching introduces one notion that has not been explored in the literature so far: the fact that, although some social contexts can be relatively stable if we consider short to moderate periods of time, our social peers are not always available at all times and spend different amounts of time in distinct social contexts.

	\section{Experimental Setup}
	\label{sec:experimental-setup}
	In this section we describe all the tools and processes necessary to produce the current research output. The experiments for our previously described simulation models were developed using the MASON~\cite{Luke2005} simulation framework written in Java. All the code for the simulation models can be found here \cite{Nunes:Software:11067}. Moreover, all the results presented in this paper were made reproducible by using the \textit{statistical computing language R} \cite{R2008} and the R package \textit{knitr} \cite{knitr2014}. Knitr is used for generating reports that contain the code, the results of its execution (plots and tables), and the textual description for such results. All the R code and the configuration files used to produce the analysed data can be found \textbf{in}...\todo{Archive the data, analysis and code used for the analysis, get a doi for it and cite it here}.

%recap
For each configuration in our experiments we performed 100 independent simulation runs. The results are thus analysed in terms of value distribution, average and standard deviation, or variance over the 100 runs. Like we described in section \ref{sec:sim_consensus}, our models are discrete-event models and each cycle agents execute 
a step in a random order : this is so that this dynamic system --our simulation model-- is not influenced by the order in which agents are executed. Simulations ends when total consensus is achieved or 2000 steps have passed.

\subsection{Social Network Models}
\label{sec:exp-setup_network_models}
The social networks used in our models are \textit{k-regular} and \textit{scale-free}. A k-regular network is a network where each node has the same number of connections. These networks are constructed by arranging the nodes in a ring and connecting each node to their next $k$ neighbours. Each network has $2k$ edges per vertex.

Scale-free networks \cite{Barabasi1999} are created with a model of preferential attachment proposed by Barab\'{a}si and colleagues \cite{Barabasi1999}. This model builds upon the perception of a common property of many large networks, a scale-free-power-law distribution of node connectivity. This feature was found to be a consequence of two generic mechanisms: networks expand continuously by the addition of new vertices, and new vertices attach preferentially to sites that are already well connected, being this mechanism more commonly known as ``preferential attachment''. In these networks, the probability $P(k)$ of two nodes being connected to each other decays as a power law, following $P(k) \sim k^{- \gamma}$.

The network instances were generated using the \textit{b-have network library} \cite{Nunes:Software:11069}. This is a Java library which allows for the manipulation of Network/Graph data structures. It also includes the more commonly used random network models from the literature.

\subsection{Measuring Network Properties}
We measure two structural properties of our networks: the \textit{average path length} and the \textit{clustering coefficient}. The \textit{average path length} measures the typical separation between two vertices in the graph. The \textit{clustering coefficient} measures the average cliquishness of the graph neighbourhood. This measure quantifies how close the neighbours of a node are to forming a clique. Duncan J. Watts and Steven Strogatz \cite{Watts1998} introduced the measure to characterise a class of complex networks called small-world networks. The networks produced by the \textit{b-have network library} \cite{Nunes:Software:11069} were exported to files and analysed using the \textit{igraph R package} \cite{igraph2006}.

@@

% Overview

\subsection{Overlapping Network Properties}
\label{sec:network_properties}
In this first analysis, we investigate the properties of different network topologies that we use in our models. 
One of the parameters in some experiments is the number of networks in which the agents interact. Adding more networks means adding more connections. 
We show that what is important is not just the the number of connections, but the properties the resulting structure displays when these networks are merged. The way we look at the network properties is as follows.

Each network is generated in such a way that the node positions are randomised. What this means is that we can have multiple networks with the same topology and yet, each node has a different neighbourhood on each network. Also, neighbourhoods do not necessarily overlap due to the node shuffling. Adding more networks to the social space is not the same as creating a single network with twice the number of connections.

\noindent Consider figure \ref{fig:network_properties_merge_2_10regular}. We created two random k-regular networks with k=10 and merged them ignoring edges from common neighbours --in the simulation models this edge agglutination is not performed, we use it here for the purposes of network analysis. The common edges are highlighted in a different colour.

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{"../analysis/pdf/ network_properties_merge_2_10regular"}
	\begin{minipage}{0.9\textwidth}
		\caption{Overlapping of two k-regular networks, each with k=10.}
		\label{fig:network_properties_merge_2_10regular}
	\end{minipage}
\end{figure}

\begin{table}
	\centering
	\begin{minipage}{0.9\textwidth}
		\caption{Properties for overlapping of two k-regular networks, each with k=10.}
		\label{tab:network_properties_merge_2_10regular}
	\end{minipage}
	\begin{tabular}{lcccc}
		& Nodes &  Edges & Clustering Coef.	  &  Avg. Path Length \\ 
		\hline  10-Regular 1 & 100 &  1,000  &  0.711 &  2.980 \\ 
		\hline  10-Regular 2 & 100 & 1,000 & 0.711 &  2.980 \\ 
		\hline  Combined & 100 & 1,790  & 0.512 &  1.638 \\ 
		\hline 20-Regular & 100 & 2,000	& 0.731	& 1.788 \\
		\hline 
	\end{tabular} 
\end{table}

The networks are drawn using the \textit{Kamada-Kawai} Layout \cite{Kamada19897} which regards the desirable ``geometric" (Euclidean) distance between two vertices in the drawing as the ``graph theoretic" distance between them in the corresponding graph. We can see by the results of the layout in figure \ref{fig:network_properties_merge_2_10regular}, that the distance between nodes in the network has decreased, this is also confirmed by the reduced average path length of the combined networks (see table \ref{tab:network_properties_merge_2_10regular}).

\begin{wrapfigure}{r}{0.45\linewidth}
	\includegraphics[width=1\linewidth]{"../analysis/pdf/ network_properties_1_20regular"}
	\begin{minipage}{0.8\linewidth}
		\caption{Single k-regular network with k=20. (Drawn using the Kamada-Kawai layout)}
		\label{fig:network_properties_1_20regular}
	\end{minipage}
\end{wrapfigure}

To illustrate our point we analysed one k-regular network with k = 20 (see figure \ref{fig:network_properties_1_20regular} --and the corresponding entry on table \ref{tab:network_properties_merge_2_10regular}). The clustering coefficient of a 20-Regular network is approximately the same as the one of a 10-regular network (higher than the combination of 10-regular networks). The average path length also decreases as there are more connections between the nodes. As we will show, these properties may vary. Since the nodes are subjected to random permutations, we can generate different overlaps which lead to distinct structural values.

Since we model the agent social space as a multitude of networks, we need to investigate the properties resulting from the merging of these networks. We do this empirically by taking the network instances used in the simulation and analysing the distribution of clustering coefficient and average path length for the different configurations.

\subsubsection{Properties of Overlapping K-Regular Networks}
\label{sec:overlapping_kreg}
We will now investigate, what kind of properties we can get from merging K-Regular networks. First, we analysed the distribution for the average path length and the clustering coefficient values over 100 network instances with 100 nodes each. Since this variation is negligible we use the average as a descriptor for different configurations of k values and number of overlapping networks. (The reader should refer to figures \ref{append_fig:network_properties_apl_kreg} and \ref{append_fig:network_properties_cc_kreg} in the appendix for the box plot preliminary analysis.)

Figure~\ref{fig:network_properties_apl_line_kreg} shows the average value for the \textit{average path length} of 100 k-regular network instances, each with 100 nodes. We can see that the \textit{average path length} changes more drastically when we go from 1 to 2 networks this is precisely the effect we can see in figure \ref{fig:network_properties_merge_2_10regular}. What we are seeing is that merging these networks at random effectively creates multiple shortcuts between points that were not connected in the original k-regular topology. Beyond this point, adding more networks does not modify this structural property in a significant way. Note that, since each network has 100 nodes, with $k=50$ the network is fully connected (hence the average path length being 1).

\begin{figure}[H]
	\centering
	\begin{subfigure}{.5\linewidth}
		\centering
		\includegraphics[width=1\linewidth]{"../analysis/pdf/ network_properties_apl_line_kreg_12345"}
		\caption{}
		\label{fig:network_properties_apl_line_kreg_12345}
	\end{subfigure}%
	\begin{subfigure}{.5\linewidth}
		\centering
		\includegraphics[width=1\linewidth]{"../analysis/pdf/ network_properties_apl_line_kreg_1020304050"}
		\caption{}
		\label{fig:network_properties_apl_line_kreg_1020304050}
	\end{subfigure}
	\begin{minipage}{0.9\textwidth}
		\vspace{0.2cm}
		\caption{Average value for the \textit{average path length} of 100 instances of overlapping k-regular networks(containing 100 nodes each) with ~k=\{1,2,3,4,5\}~(\ref{fig:network_properties_apl_line_kreg_12345}), and k= ~\{10,20,30,40,50\} (\ref{fig:network_properties_apl_line_kreg_1020304050}).}
		\label{fig:network_properties_apl_line_kreg}
	\end{minipage}
\end{figure}

\noindent Figure~\ref{fig:network_properties_cc_line_kreg} shows the results for the average \textit{clustering coefficient} of 100 instances of overlapping k-regular networks. We can see that adding multiple networks changes the clustering coefficient of the resulting structure. We can see that for $k \le 20$ the clustering coefficient drops when we add a second network. This happens because when we merge these networks, their connectivity is not enough to generate shortcuts with tightly clustered neighbourhoods. The clustering coefficient is computed by taking the average local clustering coefficient of each node in the network. An initial highly clustered k-regular network (with all the nodes having the same structure) is affected when you modify the node neighbourhoods in such a way that some are more clustered than others.

\begin{figure}[H]
	\centering
	\begin{subfigure}{.5\linewidth}
		\centering
		\includegraphics[width=1\linewidth]{"../analysis/pdf/ network_properties_cc_line_kreg_12345"}
		\caption{}
		\label{fig:network_properties_cc_line_kreg_12345}
	\end{subfigure}%
	\begin{subfigure}{.5\linewidth}
		\centering
		\includegraphics[width=1\linewidth]{"../analysis/pdf/ network_properties_cc_line_kreg_1020304050"}
		\caption{}
		\label{fig:network_properties_cc_line_kreg_1020304050}
	\end{subfigure}
	\begin{minipage}{0.9\textwidth}
		\vspace{0.2cm}
		\caption{Average value for the \textit{clustering coefficient} of 100 instances of overlapping k-regular networks(containing 100 nodes each) with ~k=\{1,2,3,4,5\}~(\ref{fig:network_properties_cc_line_kreg_12345}), and  k = \{10,20,30,40,50\} (\ref{fig:network_properties_cc_line_kreg_1020304050}).}
		\label{fig:network_properties_cc_line_kreg}
	\end{minipage}
\end{figure}

The exception is when you merge networks with $k=1$. These have a ring-like structure so there are no nodes forming a clique\footnote{A clique is a group of nodes such that every two nodes are connected by an edge.} with their neighbours.
